Spatial phylogenetics of Japanese ferns: Patterns, processes, and implications for conservation

### 2022-01-21 ms/SI.Rmd
Update MS

**`r s_figure("dendrogram")`**. Bioregion dendrograms. Tips are sites (names not shown). Dendrograms generated by clustering taxonomic (Sørensen) or phylogenetic (PhyloSor) distances between sites (see Materials and Methods). Color shows bioregion; bioregions not consisting of > `r english2(bioregion_cutoff)` grid-cells each are lumped into the "Other" category. (A), taxonomic bioregions. (B), phylogenetic bioregions. Dendrograms visualized with the `r pack("ggtree")` R package [@Yu2017].
\clearpage

@@

**`r s_figure("bioregions-env")`**. Scatterplots of grid cell membership in (A) taxonomic or (B) phylogenetic bioregions arranged by mean annual temperature and annual precipitation. Points shown with partial transparency; darker areas indicate overlapping points. Bioregions not consisting of > `r english2(bioregion_cutoff)` grid-cells each are lumped into the "Other" category.

>>>>

**`r s_figure("dendrogram")`**. Bioregion dendrograms. Tips are sites (names not shown). Dendrograms generated by clustering taxonomic (Sørensen) or phylogenetic (PhyloSor) distances between sites (see Materials and Methods). Color shows bioregion; bioregions not consisting of more than `r english2(bioregion_cutoff)` grid-cells each are lumped into the "Other" category. (A), taxonomic bioregions. (B), phylogenetic bioregions. Dendrograms visualized with the `r pack("ggtree")` R package [@Yu2017].
\clearpage

@@

**`r s_figure("bioregions-env")`**. Scatterplots of grid cell membership in (A) taxonomic or (B) phylogenetic bioregions arranged by mean annual temperature and annual precipitation. Points shown with partial transparency; darker areas indicate overlapping points. Bioregions not consisting of more than `r english2(bioregion_cutoff)` grid-cells each are lumped into the "Other" category.

### 2022-01-21 ms/manuscript.Rmd
Update MS

```{r author-page-doc, results = "asis", echo = FALSE, eval = params$doc_type == "docx"}
# Author page, version for submission to journal
cat('
Manuscript received _______; revision accepted _______.

@@

All computational analyses were carried out in R v`r getRversion()` [@RCoreTeam2020] unless otherwise stated. The R package `r pack("targets")` was used to control analysis workflow [@Landau2021].

@@

We measured functional diversity (FD) following the method of @Petchey2002.
We first calculated functional distances using Gower's method [@Gower1971] on the trait data with the `r func("gowdis")` function in the R package `r pack("FD")` [@Laliberte2010].
We weighted continuous and qualitative traits equally, weighted each binary component trait of each qualitative trait equally, and log-transformed and scaled continuous traits prior to calculating distances.
We also tested alternative weighting schemes, but these produced very similar results (data not shown).
We then used the distances to build a trait dendrogram using the `r func("hclust")` function in the R package `r pack("stats")` [@RCoreTeam2020], and used the dendogram in place of a phylogenetic tree to calculate PD and RPD as described above.
These functional analogs of PD and RPD are hereafter referred to as FD (functional diversity) and RFD (relative functional diversity).

@@

We tested whether centers of endemism have an over-representation of old (paleoendemic) or new (neoendemic) lineages using categorical analysis of neo- and paleo-endemism (CANAPE), which involves measuring PE with an alternate tree where all branch lengths are equal, then comparing the ratio between raw PE and alternative PE (RPE) [@Mishler2014].
Low RPE indicates short range-restricted branches (neoendemism) and high RPE indicates long range-restricted branches (paleoendemism).
Since all of these measures are affected by species richness, categorization of endemism is done using the rank of each observed variable relative to those calculated for the null distribution. All measures of biodiversity and CANAPE were carried out using the R package `r pack("canaper")` [@Nitta2021a].

@@

Besides habitat loss due to human activity, one major threat to ferns in Japan is herbivory by Japanese deer (*Cervus nippon*).
Although native, Japanese deer have caused extensive damage to native plant communities due to rapid population growth and range expansion since the 1970s [@Takatsuki2009].
Decrease or extirpation of fern populations due to deer herbivory has been observed in multiple sites and species across the country [@Minamitani2005; @Yahara2006; @Hattori2010].
Furthermore, 33 of 212 (15.5%) fern taxa listed as endangered (Red List status CR, EN or VU) in Japan include herbivory by deer as one of the reasons for their endangered status, and one taxon (*Arachniodes yasu-inouei* Sa. Kurata var. *angustipinnula* Seriz.) is thought to have gone extinct due to deer herbivory, among other causes [@MOJ2015; @kankyosho2022].
Although not all species of ferns are equally susceptible to deer herbivory [a small number are apparently unpalatable to deer\; @Minamitani2005], these studies show it is a clear danger to many species.
Therefore, analysis of the threat posed to fern biodiversity by deer herbivory is needed to inform future conservation policies.

>>>>

```{r author-page-doc, results = "asis", echo = FALSE, eval = params$doc_type == "doc"}
# Author page, version for submission to journal
cat('
Manuscript received _______; revision accepted _______.

@@

All computational analyses were carried out in R v`r getRversion()` [@RCoreTeam2021] unless otherwise stated. The R package `r pack("targets")` was used to control analysis workflow [@Landau2021].

@@

We measured functional diversity (FD) following the method of @Petchey2002.
We first calculated functional distances using Gower's method [@Gower1971] on the trait data with the `r func("gowdis")` function in the R package `r pack("FD")` [@Laliberte2010].
We weighted continuous and qualitative traits equally, weighted each binary component trait of each qualitative trait equally, and log-transformed and scaled continuous traits prior to calculating distances.
We also tested alternative weighting schemes, but these produced very similar results (data not shown).
We then used the distances to build a trait dendrogram using the `r func("hclust")` function in the R package `r pack("stats")` [@RCoreTeam2021], and used the dendogram in place of a phylogenetic tree to calculate PD and RPD as described above.
These functional analogs of PD and RPD are hereafter referred to as FD (functional diversity) and RFD (relative functional diversity).

@@

We tested whether centers of endemism have an over-representation of old (paleoendemic) or new (neoendemic) lineages using categorical analysis of neo- and paleo-endemism (CANAPE), which involves measuring PE with an alternate tree where all branch lengths are equal, then comparing the ratio between raw PE and alternative PE (RPE) [@Mishler2014].
Low RPE indicates short range-restricted branches (neoendemism) and high RPE indicates long range-restricted branches (paleoendemism).
Since all of these measures are affected by species richness, categorization of endemism is done using the rank of each observed variable relative to those calculated for the null distribution. All measures of biodiversity and CANAPE were carried out using the R package `r pack("canaper")` [@Nitta2022].

@@

Besides habitat loss due to human activity, one major threat to ferns in Japan is herbivory by Japanese deer (*Cervus nippon*).
Although native, Japanese deer have caused extensive damage to native plant communities due to rapid population growth and range expansion since the 1970s [@Takatsuki2009].
Decrease or extirpation of fern populations due to deer herbivory has been observed in multiple sites and species across the country [@Minamitani2005; @Yahara2006; @Hattori2010].
Furthermore, 33 of 212 (15.5%) fern taxa listed as endangered (Red List status CR, EN or VU) in Japan include herbivory by deer as one of the reasons for their endangered status [@MOJ2015; @kankyosho2022].
Although not all species of ferns are equally susceptible to deer herbivory [some are unpalatable to deer\; @Minamitani2005], these studies show it is an existential threat to many species.
Therefore, analysis of the threat posed to fern biodiversity by deer herbivory is needed to inform future conservation policies.
